Realizability-Preserving DG-IMEX Method for the Two-Moment Model of Fermion Transport \tnoteref{support}

### 2018-06-27 Documents/M1/realizableFermionicM1/sections/appendix.tex
Some reworking on Appendix C, and some changes to imex.tex

%It remains to find And the rest inequalities and the problem were left to be: \\
%Find $(x,y)$ such that
%\begin{align*}
%\text{maximize} ~ c_{sch} = &\min \{ 2x, \dfrac{1-2x(1-y)}{y-x}, 2(1-y)\}\\
% \text{under these constraints:} \qquad & x > 0, ~ 1 \geq y > 0, \\
% & x \leq y, ~\dfrac{1}{2x} \geq 1-y.
%\end{align*}
%A solution family is $x=\frac{1}{2}$, $y=\frac{1}{2} + \epsilon$ with $\epsilon \in [0, \frac{1}{2})$ and $c_{sch} = 1 - 2\epsilon$.
%The PD-ARS scheme follows.
%It's clear that $\epsilon = 0$ and $c_{sch} = 1$ is the most optimized solution.

\section{Proof of nonexistence of 3-stage Type A ...}
\label{app:noTypeA}

Proof of GSA IMEX schemes of Type~A with 3 stages and the following properties does not exist
\begin{itemize}
  \item Second-order in the streaming limit;
  \item Well-behaved in the diffusion limit;
  \item Positivity-preserving.
\end{itemize}
First of all, the GSA IMEX scheme of Type A with 3 stages have the following form:
\begin{equation*}
  \begin{array}{c | c c c}
  	      0       & 0              & 0              & 0 \\
  	\tilde{c}_{2} & \tilde{a}_{21} & 0              & 0 \\
  	\tilde{c}_{3} & \tilde{a}_{31} & \tilde{a}_{32} & 0 \\ \hline
  	              & \tilde{a}_{31} & \tilde{a}_{32} & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
    c_{1} & a_{11} & 0      & 0      \\
  	c_{2} & a_{21} & a_{22} & 0      \\
  	c_{3} & a_{31} & a_{32} & a_{33} \\ \hline
  	      & a_{31} & a_{32} & a_{33}
  \end{array}.
\end{equation*}
The consistency of the discrete system to the continuous system requires
\begin{align}
 \tilde{a}_{31} + \tilde{a}_{32} = 1, \quad \quad a_{31} + a_{32} + a_{33} = 1.
 \label{eq:1stOrderGeneral}
\end{align}
The second-order in the streaming limit requires
\begin{align}
\tilde{a}_{32}\cdot \tilde{a}_{21} =  \frac{1}{2}.
\end{align}
The diffusion accurate Eq.~\eqref{eq:diffusionCondition} requires 
\begin{align}
\tilde{a}_{21} = a_{22}, \quad -\dfrac{a_{32} \tilde{a}_{21}}{a_{22}a_{33}} + \dfrac{1}{a_{33}} =1.
\label{eq:diffusionAccurate}
\end{align}
Substituting the second equation in Eq.~\eqref{eq:1stOrderGeneral} into Eq.~\eqref{eq:diffusionAccurate} gives
\begin{align}
\tilde{a}_{21} = a_{22}, \quad a_{31} = 0.
\label{eq:a31is0}
\end{align}
Positivity-preserving requires
\begin{align*}
& a_{11} > 0, ~ a_{22} >0, ~ a_{33}>0,\\
& c_{20} = 1 - \frac{a_{21}}{a_{11}} \geq 0, ~ c_{30} = 1-\frac{a_{31}}{a_{11}} + \frac{a_{32}a_{21}}{a_{22}a_{11}} - \frac{a_{32}}{a_{22}}\geq 0, \\
& c_{21} = \frac{a_{21}}{a_{11}} \geq 0, ~ c_{31} = \frac{a_{31}}{a_{11}} - \frac{a_{32}a_{21}}{a_{22}a_{11}} \geq 0, ~ c_{32} = \frac{a_{32}}{a_{22}} \geq 0,\\
& \tilde{c}_{21} = \tilde{a}_{21} \geq 0, \tilde{c}_{31} = \tilde{a}_{31} - \frac{a_{32}\tilde{a}_{21}}{a_{22}} \geq 0, \tilde{c}_{32} = \tilde{a}_{32} \geq 0.
\end{align*}
Note that since $a_{31} = 0$, $a_{21} \geq 0$, $a_{32} \geq 0$, $a_{22} > 0$ and $a_{11} >0$, $c_{31} \geq 0$ only holds for $c_{31} = 0$.
Unfortunately, $c_{31} = 0$ will drag $c_{\mbox{\tiny Sch}} = 0$ since $c_{\mbox{\tiny Sch}} = \min \left\lbrace \dfrac{c_{21}}{\tilde{c}_{21}}, \dfrac{c_{31}}{\tilde{c}_{31}}, \dfrac{c_{32}}{\tilde{c}_{32}}\right\rbrace $.
Therefore, GSA IMEX schemes of Type~A with 3 stages and the above properties does not exist.

>>>>

\clearpage

\section{Nonexistence of a three-stage PD-IMEX Scheme of Type~A}
\label{app:noTypeA}

Here we prove that a PD-IMEX scheme (i.e., conforming to Definition~\ref{def:PD-IMEX}) of type~A does not exist.  
First, for a three-stage, GSA IMEX scheme of type~A the double Butcher tableau is
\begin{equation*}
  \begin{array}{c | c c c}
  	      0       & 0              & 0              & 0 \\
  	\tilde{c}_{2} & \tilde{a}_{21} & 0              & 0 \\
  	\tilde{c}_{3} & \tilde{a}_{31} & \tilde{a}_{32} & 0 \\ \hline
  	              & \tilde{a}_{31} & \tilde{a}_{32} & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
    c_{1} & a_{11} & 0      & 0      \\
  	c_{2} & a_{21} & a_{22} & 0      \\
  	c_{3} & a_{31} & a_{32} & a_{33} \\ \hline
  	      & a_{31} & a_{32} & a_{33}
  \end{array}.
\end{equation*}
First we consider the equality constraints.  
Consistency of the implicit coefficients and second-order accuracy in the streaming limit (Eqs.~\eqref{eq:implicitConsistency} and \eqref{eq:orderConditionsEx}, respectively) give
\begin{align}
  a_{31} + a_{32} + a_{33} = 1, \quad \tilde{a}_{31} + \tilde{a}_{32} = 1, \quad \text{and}\quad \tilde{a}_{32}\,\tilde{a}_{21} =  \f{1}{2}.
  \label{eq:1stOrderGeneral}
\end{align}
Accuracy in the diffusion limit (Eq.~\eqref{eq:diffusionCondition}) requires 
\begin{align}
  \tilde{a}_{21} = a_{22}
  \quad\text{and}\quad 
  -\f{a_{32} \tilde{a}_{21}}{a_{22}a_{33}} + \f{1}{a_{33}} =1.
\label{eq:diffusionAccurate}
\end{align}
The first constraint in Eq.~\eqref{eq:1stOrderGeneral} together with the constraints in Eq.~\eqref{eq:diffusionAccurate} gives
\begin{align}
  a_{31} = 0.
  \label{eq:a31is0}
\end{align}
The positivity-preserving property requires $a_{11}, a_{22}, a_{33}>0$ and the inequality constraints
\begin{align*}
  & c_{20} = 1 - \frac{a_{21}}{a_{11}} \geq 0, \quad c_{30} = 1-\frac{a_{31}}{a_{11}} + \frac{a_{32}a_{21}}{a_{22}a_{11}} - \frac{a_{32}}{a_{22}}\geq 0, \\
  & c_{21} = \frac{a_{21}}{a_{11}} \geq 0, \quad c_{31} = \frac{a_{31}}{a_{11}} - \frac{a_{32}a_{21}}{a_{22}a_{11}} \geq 0, \quad c_{32} = \frac{a_{32}}{a_{22}} \geq 0,\\
  & \tilde{c}_{21} = \tilde{a}_{21} \geq 0, \quad \tilde{c}_{31} = \tilde{a}_{31} - \frac{a_{32}\tilde{a}_{21}}{a_{22}} \geq 0, \quad \tilde{c}_{32} = \tilde{a}_{32} \geq 0.
\end{align*}
Since $a_{31} = 0$, we have
\begin{equation*}
  c_{31} = - \frac{a_{32}a_{21}}{a_{22}a_{11}}.  
\end{equation*}
Since $a_{21} \geq 0$, $a_{32} \geq 0$, $a_{22} > 0$ and $a_{11} >0$, we have $c_{31} \leq 0$, and the inequality constraints hold only for $c_{31} = 0$.
Unfortunately, $c_{31} = 0$ gives $c_{\mbox{\tiny Sch}} = \min \Big\{\,\f{c_{21}}{\tilde{c}_{21}}, \f{c_{31}}{\tilde{c}_{31}}, \f{c_{32}}{\tilde{c}_{32}}\Big\} = 0$.
Therefore, a three-stage PD-IMEX scheme (Definition~\ref{def:PD-IMEX}) of type~A does not exist.

### 2018-06-27 Documents/M1/realizableFermionicM1/sections/imex.tex
Reworked Appendix B
Some reworking on Appendix C, and some changes to imex.tex

Unfortunately, the correction step in Eq.~\eqref{eq:imexCorrection} deteriorates the accuracy of the IMEX scheme when applied to the moment equations in the diffusion limit ($\xi=0$, $\tau\ll 1$), where $\vect{\cK}\approx\f{1}{3}\,\cJ\,\vect{I}$.  
In the diffusion limit, to leading order in $\tau$, we then have $\vect{\cH}\approx-\f{1}{3}\,\tau\,\nabla\cJ$, which is due to a balance between the transport term and the collision term in the equation for the particle flux $\vect{\cH}$.  
The absence of the transport operator in the correction step, destroys this balance.  
We demonstrate the inferior performance of IMEX schemes with the correction step given by Eq.~\eqref{eq:imexCorrection} in the diffusion limit in Section~\ref{sec:smoothProblems}.  
We have also implemented and tested one of the IMEX schemes in Chertock et al. \cite{chertock_etal_2015}, where the transport operator is included in the correction step, and found it to perform very well in the diffusion limit.  
However, we have not been able to prove the realizability-preserving property with this approach without invoking a too severe time step restriction.  
We therefore proceed to design positivity-preserving IMEX schemes, without the correction step, that perform better in the diffusion limit.  
We limit the scope to IMEX schemes of type~ARS.  
(It can be shown that IMEX schemes of type~A conforming to Definition~\ref{def:DIMEXP} below do not exist; cf. \ref{app:noTypeA}.)

@@

@@

  \label{def:DIMEXP}

@@

Fortunately, IMEX schemes satisfying these properties are easy to find, and we provide an example with $s=3$ (two implicit solves) in \ref{app:butcherTables} (scheme PD-ARS; see \ref{app:PD-ARS} for details).  

>>>>

Unfortunately, the correction step in Eq.~\eqref{eq:imexCorrection} deteriorates the accuracy of the IMEX scheme when applied to the moment equations in the diffusion limit ($\xi=0$, $\tau\ll 1$), where $\vect{\cK}\approx\f{1}{3}\,\cJ\,\vect{I}$.  
In the diffusion limit, to leading order in $\tau$, we then have $\vect{\cH}\approx-\f{1}{3}\,\tau\,\nabla\cJ$, which is due to a balance between the transport term and the collision term in the equation for the particle flux $\vect{\cH}$.  
The absence of the transport operator in the correction step, destroys this balance.  
We demonstrate the inferior performance of IMEX schemes with the correction step given by Eq.~\eqref{eq:imexCorrection} in the diffusion limit in Section~\ref{sec:smoothProblems}.  
We have also implemented and tested one of the IMEX schemes in Chertock et al. \cite{chertock_etal_2015}, where the transport operator is included in the correction step, and found it to perform very well in the diffusion limit.  
However, we have not been able to prove the realizability-preserving property with this approach without invoking a too severe time step restriction.  
We therefore proceed to design positivity-preserving IMEX schemes, without the correction step, that perform better in the diffusion limit.  
We limit the scope to IMEX schemes of type~ARS.  
(It can be shown that IMEX schemes of type~A conforming to Definition~\ref{def:PD-IMEX} below do not exist; cf. \ref{app:noTypeA}.)

@@

      \label{eq:implicitConsistency}

@@

  \label{def:PD-IMEX}

@@

Fortunately, IMEX schemes of type~ARS satisfying these properties are easy to find, and we provide an example with $s=3$ (two implicit solves) in \ref{app:butcherTables} (scheme PD-ARS; see \ref{app:PD-ARS} for details).  
